### 2019-02-14 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Trying to finish chapter 4 as quickly as possible.

% ----------------------------------------------------------------------
\subsection{Retransmissions in an ALOHA like protocol}\label{sub:43:introduction}

@@

% \vspace*{5pt}
The rest of the section is organized as follows.
First the system model is introduced in Section~\ref{sub:43:model},
% A formal description of the MAB learning algorithms is given in Section~\ref{sub:43:motivations},
and our motivations are exposed in Section~\ref{sub:43:MABalgo}.
The proposed \UCB-based heuristics are presented in Section~\ref{sub:43:heuristics}, while the corresponding numerical results are shown in Section~\ref{sub:43:numExp}.
Finally, some conclusions are drawn in Section~\ref{sub:43:conclusion}.

% ----------------------------------------------------------------------
\subsection{System model}\label{sub:43:model}

@@

% ----------------------------------------------------------------------
\subsection{Motivations for the proposed approach} \label{sub:43:motivations}

@@

% --- --- --- --- --- --- --- --- --- ---
\paragraph{Behavior analysis of $p_{c}$ and $p_{c1}$}\label{sub:numericalValidationPC1PC}

@@

%-----------------------------------------------------------------
\subsection{A well-known MAB Algorithm: \UCB{}}\label{sub:43:MABalgo}

Without loss of generality, we have adopted a well-studied stochastic MAB learning algorithm, where the reward distributions are unknown and assumed to be independent and identically distributed (i.i.d). The arms model the channels denoted as $C(t) \in \llbracket 1, K \rrbracket$, and the players, the dynamic devices, learn the distributions to be able to progressively focus on the best arm, \ie, the arm with largest mean representing the mean availability of a given channel $k$.

Before presenting our proposed heuristics, we describe a \UCB{} bandit algorithm \cite{Auer02}. It has reported to be efficient, while featuring a low complexity for its implementation. For this reason, it has been employed for IoT applications \cite{Bonnefoi17}, and we employ this approach to develop our proposals.

% % --- --- --- --- --- --- --- --- --- ---
\paragraph{The \UCB{} algorithm}\label{sub:algoUCB}

A first approach is to only use an empirical mean estimator of the rewards in every channel, and select the channel with highest estimated mean at every time step; but this greedy approach is known to fail dramatically \cite{Auer02}.
Indeed, with this policy, the selection of arms depends too much on the first draws: if the first transmission in one channel fails and the first one on other channels succeeds, the device will \emph{never} use the first channel again, even if it is the best one (\ie, the most available, in average).

Rather than relying on the empirical mean reward, \UCB{} algorithms instead use a \emph{confidence interval} on the unknown mean $\mu_k$ of each arm, which can be viewed as adding a ``bonus'' exploration to the empirical mean. They follow the ``\emph{optimism-in-face-of-uncertainty}'' principle: at each step, they play according to the best model, as the statistically best possible arm (\ie, the highest \UCB{}) is selected.

@@

% ----------------------------------------------------------------------
\subsection{Heuristics to (try to) learn how to retransmit efficiently}\label{sub:43:heuristics}

@@

% --- --- --- --- --- --- --- --- --- ---
\paragraph{Uniform random retransmission (``Random'')}\label{sub:UCBthenRandom}

@@

% --- --- --- --- --- --- --- --- --- ---
\paragraph{\UCB{} for retransmission (``Only \UCB{}'')}\label{sub:TwoUCB}

@@

% --- --- --- --- --- --- --- --- --- ---
\paragraph{$K$ different {\UCB}s for retransmission (``$K$ \UCB{}'')}\label{sub:UCBthenKp1}

@@

Although, this approach increases the complexity and store requirements (of order $\mathcal{O}(K^2)$).
For our LPWA networks of interest, such as LoRaWAN, the cost of its implementation is still affordable, since a small number of channels is used.
For instance, for $K=4$ channels,
the memory to store $K+1=5$ algorithms is of the order of the requirements to storing one.
% For other networks this heuristic could not be practical, for instance  Sigfox consist in a large number of very narrow-band channels.

%\footnote{~The storage requirements and time complexity is now quadratic in $K$, and as such we no longer consider this heuristic to be a practical proposal in some LPWA networks, as for instance Sigfox networks consist in a large number of very narrow-band channels. But for LoRaWAN networks with $K=4$, storing $K+1=5$ algorithms does not cost much more than storing $2$.}

@@

% --- --- --- --- --- --- --- --- --- ---
\paragraph{Delayed \UCB{} for retransmission (``Delayed \UCB{}'')}\label{sub:UCBwithDelay}

@@

The idea behind this delay is to allow the first stage \UCB{} to start learning the best channel, before starting the second stage \UCB{} (see details in Algorithm~\ref{algo:43:UCBwithDelay}).
The number of transmissions to wait before applying the second algorithm is denoted by $\Delta$, it has to be fixed before-hand.

@@

%\footnote{~Choosing the value of $\Delta$ could be done by extensive benchmarks but such approach goes against the reinforcement learning idea: an heuristic should work against any problem, without the need to be able to simulate the problem in advance in order to find a good value of some internal parameter. As such, we only consider a delay of $\Delta=100$.}.

@@

% ----------------------------------------------------------------------
\subsection{Simulations to compare our heuristics}\label{sub:43:numExp}

@@

% ----------------------------------------------------------------------
\subsection{Conclusion about this second model}\label{sub:43:conclusion}

In this section, we presented a retransmission model of LPWA networks based on an ALOHA protocol, slotted both in time and frequency, in which dynamic IoT devices can use machine learning algorithms, to improve their Packet Loss Ratio (PLR) when accessing the network.
The main novelty of this model is to address the packet retransmissions upon radio collision, by using a Multi-Armed Bandit framework.
We presented and evaluated several learning heuristic that try to learn how to transmit and retransmit in a smarter way, by using the \UCB{} algorithm for channel selection for first transmission, and different proposals based on \UCB{} for the retransmissions upon collisions.

We showed that incorporating learning for the transmission is needed to achieve optimal performance, with significant gain in terms of successful transmission rate in networks with a large number of devices (up-to $30\%$ in the example network).
Our empirical simulations show that each of our proposed heuristic outperforms a naive random access scheme.
Surprisingly, the main take-away message is that a simple \UCB{} learning approach, that retransmit in the same channel, turns out to perform as well as more complicated heuristics.

\paragraph{Future works}
%
The utility and impact of the proposed approaches for LPWA networks motivates us to address several subjects as future works. Among them, the non-stationarity of the channel occupancy caused by the learning policy employed by the IoT devices.
%
For that end, modifications of MAB algorithms have been proposed, such as Sliding-Window-\UCB{} or Discounted-\UCB{} \cite{Garivier08}
or more recently M-\UCB{} \cite{CaoZhenKvetonXie18},
% or more recently GLR-\UCB{} \cite{Besson2019GLRT},
that nevertheless have not been explored for the targeted problem.
Chapter~\ref{chapter:6} is focusing on this direction, but we did not have enough time to explore the possible applications of MAB policies designed for non-stationary problems to the model with retransmissions presented in this section.

In order to validate our results in a realistic experimental setting and not only with simulations, future works include a hardware implementation of the analyzed models to complete our recent works \cite{modiDemo2016,Besson2019WCNC}.
Note that Julio César Mango-Vasquez is currently working on this direction, for the EPHYL project, in collaboration with Carlos Faouzi Bader at IETR and CentraleSupélec campus of Rennes, and Christophe Moy at IETR and University Rennes 1.
%
A hardware demonstrator could be also benefit to study other settings by removing some hypotheses, for instance by studying a similar model in non-slotted time.

>>>>

% ----------------------------------------------------------------------
\subsection{Retransmissions in an ALOHA like protocol}
\label{sub:43:introduction}
% ----------------------------------------------------------------------

@@

\paragraph{Outline}

The rest of the Section is organized as follows.
First the system model is introduced in Section~\ref{sub:43:model},
% A formal description of the MAB learning algorithms is given in Section~\ref{sub:43:motivations},
and our motivations are exposed in Section~\ref{sub:43:MABalgo}.
The proposed \UCB-based heuristics are presented in Section~\ref{sub:43:heuristics}, while the corresponding numerical results are shown in Section~\ref{sub:43:numExp}.
Finally, some conclusions are drawn in Section~\ref{sub:43:conclusion}.

% ----------------------------------------------------------------------
\subsection{Extending our model with retransmissions}
\label{sub:43:model}
% ----------------------------------------------------------------------

@@

% ----------------------------------------------------------------------
\subsection{Motivations for the proposed approach}
\label{sub:43:motivations}
% ----------------------------------------------------------------------

@@

% --- --- --- --- --- --- --- --- --- ---
\paragraph{Behavior analysis of $p_{c}$ and $p_{c1}$}\label{sub:43:numericalValidationPC1PC}

@@

\TODOL{Remove this part or compress}

%-----------------------------------------------------------------
\subsection{The \UCB{} algorithm as a building block for different heuristics}
\label{sub:43:MABalgo}
% ----------------------------------------------------------------------

Before presenting our proposed heuristics, we remind here the \UCB{} bandit algorithm \cite{Auer02}.
We include it again to facilitate the understanding of the heuristics, which use \UCB{} as a building block.

% % --- --- --- --- --- --- --- --- --- ---
\paragraph{The \UCB{} algorithm}\label{sub:43:algoUCB}

@@

% ----------------------------------------------------------------------
\subsection{Heuristics to (try to) learn how to retransmit efficiently}
\label{sub:43:heuristics}
% ----------------------------------------------------------------------

@@

% --- --- --- --- --- --- --- --- --- ---
\paragraph{Uniform random retransmission (``Random'')}\label{sub:43:UCBthenRandom}

@@

% --- --- --- --- --- --- --- --- --- ---
\paragraph{\UCB{} for retransmission (``Only \UCB{}'')}\label{sub:43:TwoUCB}

@@

% --- --- --- --- --- --- --- --- --- ---
\paragraph{$K$ different {\UCB}s for retransmission (``$K$ \UCB{}'')}\label{sub:43:UCBthenKp1}

@@

Although, this approach increases the complexity and store requirements (of order $\mathcal{O}(K^2)$).
For our LPWA networks of interest, such as LoRaWAN, the cost of its implementation is still affordable, since a small number of channels is used.
For instance, for $K=4$ channels,
the memory to store $K+1=5$ algorithms is of the order of the requirements to storing one.

Note that for other networks this heuristic could not be practical, for instance Sigfox consist in a large number of very narrow-band channels.
The storage requirements and time complexity is now quadratic in $K$, and as such we no longer consider this heuristic to be a practical proposal in some LPWA networks, as for instance Sigfox networks consist in a large number of very narrow-band channels. But for LoRaWAN networks with $K=4$, storing $K+1=5$ algorithms does not cost much more than storing $2$.

@@

% --- --- --- --- --- --- --- --- --- ---
\paragraph{Delayed \UCB{} for retransmission (``Delayed \UCB{}'')}\label{sub:43:UCBwithDelay}

@@

The idea behind this delay is to allow the first stage \UCB{} to start learning the best channel, before starting the second stage \UCB{} (see details in Algorithm~\ref{algo:43:UCBwithDelay}).
The number of transmissions to wait before applying the second algorithm is denoted by $\Delta$, it has to be fixed before-hand\footnote{~Choosing the value of $\Delta$ could be done by extensive benchmarks but such approach goes against the reinforcement learning idea: an heuristic should work against any problem, without the need to be able to simulate the problem in advance in order to find a good value of some internal parameter. As such, we only consider a delay of $\Delta=100$.}.

@@

@@

% ----------------------------------------------------------------------
\subsection{Simulations to compare our heuristics}
\label{sub:43:numExp}
% ----------------------------------------------------------------------

@@

### 2019-02-14 2-Chapters/4-Chapter/chapter4.tex
Trying to finish chapter 4 as quickly as possible.

\chapter{More Efficient Spectrum Usage of IoT Networks with Selfish MAB Learning}
\label{chapter:4}
\minitoc

@@

\TODOL{Write a nice introduction for this chapter.}

% ----------------------------------------------------------------------------
\section{Motivations for Selfish MAB Learning for IoT Networks}
\label{sec:4:motivations}

@@

\TODOL{Present outline of the chapter.}

% \TODOL{Give list of publications associated with this chapter.}
\paragraph{Publications}
This Chapter is based on the following publications: \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.

% ----------------------------------------------------------------------------
\section{Selfish learning for many dynamic devices with low activation probabilities in an IoT network}
\label{sec:4:firstModel}

@@

% ----------------------------------------------------------------------------
\section{Test-bed implementation of our model for real-world validation}
\label{sec:4:gnuradio}

@@

% ----------------------------------------------------------------------------
\section{Extension of our model to account for retransmissions}
\label{sec:4:retransmissions}

@@

% ----------------------------------------------------------------------------
\section{Conclusion of Chapter~\ref{chapter:4}}
\label{sec:4:conclusion}

\TODOL{Write a conclusion, and future works}

\subsection{Summary of the first model}\label{sub:41:conclusion}

@@

\subsection{Conclusions taken from the demonstration}
\label{sec:42:conclusionFromDemonstration}

@@

\subsection{Future works}
Future works include...

\paragraph{Possible extensions of the first model}
%
The first model presented in Section~\ref{sec:4:firstModel} could easily be generalized with two probabilities $p_S$ and $p_D$, if static and dynamic devices had different transmission pattern, and less easily with one probability per device. Also, other emission pattern could be considered, instead of a Bernoulli process for each user.
We prefer to consider that all devices have the same probability to transmit to keep the notation as simple as possible.

@@

\paragraph{Extensions of the demonstration}
Possible future extensions of our demonstration include the following points.
We could consider more dynamic objects (\emph{e.g.}, $100$) but it would either cost more in terms of equipment, or in terms of software engineering to simulation more objects with the same card.
We could also implement a real-world IoT communication protocol (like the LoRaWAN standard), which we prefer not to do as it would cost a significant effort of development.
Finally, we could also study the interference in case of other gateways located nearby, and this could be done without needing new hardware.
%
% Pub pour notre autre article
% We are also interested in studying the possible gain of using a learning step when the transmission model follows ALOHA-like retransmissions, and this is presented in the next Section~\ref{sec:4:retransmissions}.

% ----------------------------------------------------------------------------
\section{Appendix for Chapter~\ref{chapter:4}}
\label{sec:4:appendix}

% \TODOL{Include here additional results (proof in Section from CROWNCOM ?)}
\input{2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted_appendix.tex}

>>>>

% ----------------------------------------------------------------------
\chapter{More Efficient Spectrum Usage of IoT Networks with Selfish MAB Learning}
\label{chapter:4}
\minitoc
% ----------------------------------------------------------------------

@@

% ----------------------------------------------------------------------------
\section{Motivations for Selfish MAB Learning for IoT Networks}
\label{sec:4:motivations}
% ----------------------------------------------------------------------

@@

% \TODOL{Present outline of the chapter.}

\paragraph{Publications}

This Chapter is based on the following publications: \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.

% ----------------------------------------------------------------------------
\section{Selfish learning for many dynamic devices with low activation probabilities in an IoT network}
\label{sec:4:firstModel}
% ----------------------------------------------------------------------

@@

% ----------------------------------------------------------------------------
\section{Test-bed implementation of our model for real-world validation}
\label{sec:4:gnuradio}
% ----------------------------------------------------------------------

@@

% ----------------------------------------------------------------------------
\section{Extension of our model to account for retransmissions}
\label{sec:4:retransmissions}
% ----------------------------------------------------------------------

@@

% ----------------------------------------------------------------------------
\section{Conclusion of Chapter~\ref{chapter:4}}
\label{sec:4:conclusion}
% ----------------------------------------------------------------------

We give in this Section conclusions about the two models presented above and about our demonstration, as well as many directions of future works, some of which are currently being studied by researchers of our team.

% ----------------------------------------------------------------------
\paragraph{Summary of the first model}
\label{sub:41:conclusion}

@@

% ----------------------------------------------------------------------
\paragraph{Conclusions taken from the demonstration}
\label{sub:42:conclusionFromDemonstration}

@@

% ----------------------------------------------------------------------
\paragraph{Conclusion about this second model}
\label{sub:43:conclusion}

In Section~\ref{sec:4:retransmissions}, we presented an extension of our model of LPWA networks based on an ALOHA protocol, slotted both in time and frequency, in which dynamic IoT devices can use machine learning algorithms, to improve their Packet Loss Ratio (PLR) when accessing the network.
The main novelty of this model is to address the packet retransmissions upon radio collision, by using a Multi-Armed Bandit framework.
We presented and evaluated several learning heuristic that try to learn how to transmit and retransmit in a smarter way, by using the \UCB{} algorithm for channel selection for first transmission, and different proposals based on \UCB{} for the retransmissions upon collisions.

We showed that incorporating learning for the transmission is needed to achieve optimal performance, with significant gain in terms of successful transmission rate in networks with a large number of devices (up-to $30\%$ in the example network).
Our empirical simulations show that each of our proposed heuristic outperforms a naive random access scheme.
Surprisingly, the main take-away message is that a simple \UCB{} learning approach, that retransmit in the same channel, turns out to perform as well as more complicated heuristics.

% ----------------------------------------------------------------------
\paragraph{Future works}
\label{sub:4:futureWorks}

Future works related to this Chapter include the following directions.

\paragraph{Possible extensions of the first model}

The first model presented in Section~\ref{sec:4:firstModel} could easily be generalized with two probabilities $p_S$ and $p_D$, if static and dynamic devices had different transmission pattern, and less easily with one probability per device. Also, other emission pattern could be considered, instead of a Bernoulli process for each user.
We prefer to consider that all devices have the same probability to transmit to keep the notation as simple as possible.

@@

\paragraph{Extensions of the demonstration}

Possible future extensions of our demonstration include the following points.
We could consider more dynamic objects (\emph{e.g.}, $100$) but it would either cost more in terms of equipment, or in terms of software engineering to simulation more objects with the same card.
We could also implement a real-world IoT communication protocol (like the LoRaWAN standard), which we prefer not to do as it would cost a significant effort of development.
Finally, we could also study the interference in case of other gateways located nearby, and this could be done without needing new hardware.
%
% Pub pour notre autre article
% We are also interested in studying the possible gain of using a learning step when the transmission model follows ALOHA-like retransmissions, and this is presented in the next Section~\ref{sec:4:retransmissions}.

\paragraph{Possible extensions for the second model}

The utility and impact of the proposed approaches for LPWA networks motivates us to address several subjects as future works. Among them, the non-stationarity of the channel occupancy caused by the learning policy employed by the IoT devices.
%
For that end, modifications of MAB algorithms have been proposed, such as Sliding-Window-\UCB{} or Discounted-\UCB{} \cite{Garivier11UCBDiscount}
or more recently M-\UCB{} \cite{CaoZhenKvetonXie18},
or more recently GLR-\UCB{} \cite{Besson2019GLRT} which is presented in Chapter~\ref{chapter:6},
that nevertheless have not been explored for the targeted problem.
Chapter~\ref{chapter:6} is focusing on this direction, but we did not have enough time to explore the possible applications of MAB policies designed for non-stationary problems to the model with retransmissions presented in this section.

In order to validate our results in a realistic experimental setting and not only with simulations, future works include a hardware implementation of the analyzed models to complete our demonstration \cite{Besson2019WCNC}.
Note that Julio César Mango-Vasquez is currently working on this direction, for the EPHYL project, in collaboration with Carlos Faouzi Bader at IETR and CentraleSupélec campus of Rennes, and Christophe Moy at IETR and University Rennes 1.
%
A hardware demonstrator could be also benefit to study other settings by removing some hypotheses, for instance by studying a similar model in non-slotted time.

% ----------------------------------------------------------------------------
\section{Appendix for Chapter~\ref{chapter:4}}
\label{sec:4:appendix}
% ----------------------------------------------------------------------------

We include here a missing proof from Section~\ref{sec:4:firstModel}.

\input{2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted_appendix.tex}
